---
title: Descriptive Statistics and Regression exam  (2023/03/23)
lang: en
date: 2023-03-23
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(kableExtra)
```

:::{#exr-1}
```{r, echo=F, results= "hide"}
## DATA GENERATION
z <- factor(rep(c("SS","AP","NT","SB"), c(18, 24, 15, 3)), levels = c("SS","AP","NT","SB"))
x <- rep(c(2.5, 6, 8, 9.5), c(18, 24, 15, 3))
df <- data.frame(z)
y <- 10 * x + 5

## QUESTION/ANSWER GENERATION
# Tamaño de la muestra
n <- length(x)
# Median
me <- 5+2*0.2/0.4
# Cumulative frequency for 8
F8 <- 0.7 + (0.95-0.7) / (9-7)
N8 <- F8 * n
# Mean
mx <- mean(x)
my <- mean(y)
# Variance
s2x <- sum(x^2)/n-mx^2
s2y <- sum(y^2)/n-my^2
# Std. deviation
sx <- sqrt(s2x)
sy <- sqrt(s2y)
# Coef. variation
cvx <- sx/mx
cvy <- sy/my
# Coef. skewness
g1x <- sum((x-mx)^3)/(n*sx^3)
g1y <- sum((y-my)^3)/(n*sy^3)
# Coef. kurtosis
g2x <- sum((x-mx)^4)/(n*sx^4)-3
g2y <- sum((y-my)^4)/(n*sy^4)-3
```


The chart below shows the percentage of grades in a Statistic course with 60 students. 

```{r pie-chart-scores, results = "hide", fig.height = 5, fig.width = 6, fig.path = "img/exam-2023-03-23/", fig.cap = ""}
library(lessR)
# Pie chart of grades
PieChart(z, hole = 0, values = "%", data = df, main = "")
```

a. Plot the ogive of the score, assuming the following correspondence between grades and scores

$$
\begin{array}{lc}
  \mbox{Grade} & \mbox{Score}\\
  \mbox{SS} & [0, 5)\\
  \mbox{AP} & [5, 7)\\
  \mbox{NT} & [7,9)\\
  \mbox{SB} & [9,10]
\end{array}
$$

a. Compute the median and interpret it.

a. How many students got a score greater than 8?

a. Study the dispersion of the distribution.

a. Study the skewness of the distribution. Is it normal?

a. If we apply the transformation $y=10x+5$ to the scores, how changes the representativeness of the mean. And the skewness?

Use the following sums for the computations ($X$ = Score):  
$\sum x_in_i=`r round(sum(x), 2)`$, $\sum x_i^2n_i=`r round(sum(x^2), 4)`$, $\sum (x_i-\bar x)^3n_i=`r round(sum((x-mx)^3),2)`$ and $\sum (x_i-\bar x)^4n_i=`r round(sum((x-mx)^4),2)`$.

:::{.callout-tip collapse="true"}
## Solution
a. Ogive 

```{r ogive-scores, results = "hide", fig.height = 5, fig.width = 6, fig.path = "img/exam-2023-03-23/", fig.cap = ""}
color1 <- "#00BFC4"
plot(c(0, 5, 7, 9, 10), c(0, 0.3, 0.7, 0.95, 1), type = "l", col = color1, xlab = "Score", ylab = "Cumulative relative frequency", axes = F)
axis(1, at = 0:10)
axis(2, at = seq(0,1,0.1))
box()
```

b. $Me = `r me`$ points.

c. $N(8) = `r N8`$ students.

d. $\bar x = `r round(mx, 4)`$ points, $s_x^2=`r round(s2x,4)`$ points$^2$, $s_x=`r round(sx,4)`$ points and $cv_x=`r round(cvx,4)`$. Thus, there is a moderate dispersion with respect to the mean.

e. $g_1 = `r round(g1x,4)`$ and therefore the distribution is a little bit left skewed.

f. $\bar y = `r round(my, 4)`$ points, $s_y^2=`r round(s2y,4)`$ points$^2$, $s_y=`r round(sy,4)`$ points and $cv_y=`r round(cvy,4)`$. As $cv_y < cv_x$ the representativeness of the mean increases. As the slope of the linear transformation is positive, the skewness does not change.
:::
:::

:::{#exr-2}
```{r, echo=F, results= "hide"}
## DATA GENERATION
set.seed(12)
x <- round(runif(40, 20, 45))
y <- 38 + x/20 + round(rnorm(40),1)

## QUESTION/ANSWER GENERATION
# colors
col1 <- "#05A1E6"
# sample size
n <- length(x)
# means
mx <- mean(x)
mlogx <- mean(log(x))
my <- mean(y)
mlogy <- mean(log(y))
# variances
s2x <- sum(x^2)/n-mx^2
s2logx <- sum(log(x)^2)/n-mlogx^2
s2y <- sum(y^2)/n-my^2
s2logy <- sum(log(y)^2)/n-mlogy^2
# Covariances
sxy <- sum(x * y) / n - mx * my
sxlogy <- sum(x*log(y))/n-mx*mlogy
slogxy <- sum(log(x)*y)/n-mlogx*my
# Determination coeff
r2xy <-  sxy^2 / (s2x * s2y)
r2xlogy <- sxlogy^2/(s2x*s2logy)
r2logxy <- slogxy^2/(s2logx*s2y)
# Linear regression model
reglin <- lm(y ~ x)
# Exponential regression model y on x
regexp <- lm(log(x)~y)
# Logarithmic regression model x on y
reglog <- lm(y~log(x))
```

A study tries to determine if there is a relation between the gestation time (in weeks) and the age of the mother (in years). A sample of 40 mothers was taken and the sums below summarize the results ($X$=Age and $Y$=Gestation time):

$\sum x_i=`r round(sum(x), 4)`$ years, $\sum \log(x_i)=`r round(sum(log(x)), 4)`$ log(years), $\sum y_j=`r round(sum(y), 4)`$ weeks, $\sum \log(y_j)=`r round(sum(log(y)), 4)`$ log(weeks),  
$\sum x_i^2=`r round(sum(x^2), 4)`$ years$^2$, $\sum \log(x_i)^2=`r round(sum(log(x)^2), 4)`$ log(years)$^2$, $\sum y_j^2=`r round(sum(y^2), 4)`$ weeks$^2$, $\sum \log(y_j)^2=`r round(sum(log(y)^2), 4)`$ log(weeks)$^2$,  
$\sum x_iy_j=`r round(sum(x*y), 4)`$ years$\cdot$weeks, $\sum x_i\log(y_j)=`r round(sum(x*log(y)), 4)`$ years$\cdot$log(weeks), $\sum \log(x_i)y_j=`r round(sum(log(x)*y), 4)`$ log(years)$\cdot$weeks, $\sum \log(x_i)\log(y_j)=`r round(sum(log(x)*log(y)), 4)`$ log(years)$\cdot$log(weeks).

a. Which regression models, linear, exponential or logarithmic, explains better the relation between the age and the gestation time? 

a. Use the best model to predict the gestation time for a mother 45 years old. Is this prediction reliable?

a. According to the linear model, how much increases or decreases the gestation time for every year of the mother?

:::{.callout-tip collapse="true"}
## Solution
a. Linear model: $\overline{x}=`r round(mx,4)`$ years, $s_x^2=`r round(s2x,4)`$ years$^2$.  
$\bar y=`r round(my,4)`$ weeks, $s_y^2=`r round(s2y,4)`$ weeks$^2$.  
$s_{xy}=`r round(sxy,4)`$ years$\cdot$weeks.  
$r^2 = `r round(r2xy, 4)`$. 

    Exponential model: $\overline{\ln(y)} = `r round(mlogy, 4)`$ ln(weeks), $s_{\ln(y)}^2 = `r round(s2logy, 4)`$ ln(weeks)$^2$  
    $s_{x\ln(y)} = `r round(sxlogy, 4)`$ years$\cdot\ln$(weeks).  
    $r^2 = `r round(r2xlogy, 4)`$.  

    Logarithmic model: $\overline{\ln(x)} = `r round(mlogx, 4)`$ ln(years), $s_{\ln(x)}^2 = `r round(s2logx, 4)`$  ln(years)$^2$  
    $s_{\ln(x)y} = `r round(slogxy, 4)`$ ln(years)weeks.  
    $r^2 = `r round(r2logxy, 4)`$.

    As the linear coefficient of determination is greater, the linear model explains better the relation between de gestation time and the age of the mother. 

a. 
Linear regression model of $Y$ on $X$: $y = `r round(reglin$coefficients[1], 4)` + `r round(reglin$coefficients[2], 4)` x$.  
Predictions: $y(45) = `r round(reglin$coefficients[1] + reglin$coefficients[2] * 45, 4)`$ weeks.  
The predictions are not reliable because the coefficient of determination is pretty low.

a. Regression coefficient of $Y$ on $X$: $b_{yx} = `r round(reglin$coefficients[2], 4)`$ weeks/year. The gestation time increases $`r round(reglin$coefficients[2], 4)`$ weeks per year.
:::
:::