---
title: 2022-05-06<br>Probability and Random Variables Exam
date: 2022-05-06
lang: en
---

:::{#exr-1}
```{r echo=FALSE, results="hide"}
## DATA GENERATION
pC <- 0.08
pHsiC <- 0.040
pCsiH <- 0.1
## QUESTION/ANSWER GENERATION
pCH <- 0.0032
pCnoH <- 0.0768
pnoCH <- 0.0288
pnoCnoH <- 0.8912
rC <- pCH / (pCH + pCnoH)
rnoC <- pnoCH / (pnoCH + pnoCnoH)
rr <- rC / rnoC
oC <- pCH / pCnoH
onoC <- pnoCH / pnoCnoH
or <- oC / onoC
```

8% of people in a population consume cocaine. It is also known that 4%: of people who consume cocaine have a heart attack and 10%: of people who have a heart attack consume cocaine.


a. Construct the probability tree for the random experiment of drawing a random person from the population and measuring if he or she consumes cocaine and if he or she has a heart attack.

a. Compute the probability that a random person of the population does not consume cocaine and does not have a heart attack.

a. Are the events of consuming cocaine and having a heart attack dependent?

a. Compute the relative risk and the odds ratio of suffering a heart attack consuming cocaine. Which association measure is more suitable for this study? Interpret it.

:::{.callout-tip collapse="true"}
## Solution
Let $C$ the event of consuming cocaine and $H$ the event of having a heart attack.

a.

:::{.content-visible when-format="html"}
![Probability tree](img/exam-2022-05-06/probability-tree-cocaine.svg)
:::

:::{.content-visible unless-format="html"}
![Probability tree](img/exam-2022-05-06/probability-tree-cocaine.pdf)
:::

a. $P(\overline C\cap \overline H)=`r round(pnoCnoH, 4)`$.

a. The events are dependent as $P(C)=`r pC`\neq P(C|H)=`r pCsiH`$.

a. $RR(H)=`r round(rr, 4)`$ and $OR(H)=`r round(or, 4)`$. The odds ratio is more suitable as the study is retrospective. That means that the odds of having a heart attack is $`r round(or, 4)`$ times greater if a person consumes cocaine.
:::
:::

:::{#exr-2}
```{r echo=FALSE, results="hide"}
## DATA GENERATION
lambda1 <- 3
lambda2 <- 12
n <- 10
p <- 0.05
## QUESTION/ANSWER GENERATION
p1 <- ppois(4, lambda1, lower.tail = F)
p2 <- ppois(5, lambda2)
p3 <- pbinom(0, 10, p2, lower.tail = F)
```

A basketball player scores 12 points per game on average.


a. What is the probability that the player scores more than 4 points in a quarter?

a. If the player plays 10 games in a league, what is the probability of scoring less than 6 points in some game?

:::{.callout-tip collapse="true"}
## Solution

a. Let $X$ be the points scored in a quarter by the player. Then $X\sim P(3)$, and $P(X>4)=`r round(p1, 4)`$.

a. Let $Y$ be the number of points scored in a game by the player. Then $Y\sim P(12)$ and $P(Y<6)=`r round(p2,4)`$.  
Let $Z$ be the number of games with less than 6 points scored by the player. Then $Z\sim B(10, `r round(p2,4)`)$, and $P(Z>0)=`r round(p3,4)`$.
:::
:::

:::{#exr-3}
```{r echo=FALSE, results="hide"}
## DATA GENERATION
m <- 174
s <- 51.938
m2 <- 350

## QUESTION/ANSWER GENERATION
spe <- pnorm(300, m, s)
sen <- pnorm(300, m2, s, lower.tail = F)
pre <- 0.08
truepos <- pre * sen
falseneg <- pre * (1 - sen)
falsepos <- (1 - pre) * (1 - spe)
trueneg <- (1 - pre) * (spe)
ppv = truepos / (truepos + falsepos)
npv = trueneg / (falseneg + trueneg)
```

The creatine phosphokinase (CPK3) is an enzyme in the body that causes the phosphorylation of creatine. This enzyme is found in the skeletal muscle and can be measured in a blood analysis. The concentration of CPK3 in blood is normally distributed, and the interval centered at the mean with the reference values, that accumulates 99%: of the population, ranges from 40 to 308 IU/L in healthy adult males.
 
a. Compute the mean and the standard deviation of the concentration of CPK3 in healthy males.  
Note: If you are not able to compute the standard deviation, use $\sigma = 50$ UI/L for the next parts.

a. A diagnostic test to detect muscular dystrophy gives a negative outcome when the concentration of CPK3 is below 300 UI/L. Compute the specificity of the test.

a. If the concentration of CPK3 in people with muscular dystrophy also follows a normal distribution with mean 350 IU/L and the same standard deviation, what is the sensitivity of the test?

a. Compute the predictive values of the test and interpret them assuming that the muscular dystrophy prevalence is 8%.

:::{.callout-tip collapse="true"}
## Solution

a. $\mu = `r m`$ IU/L and $\sigma = `r round(s,4)`$ IU/L.

a. Specificity = $`r round(spe,4)`$.

a. Sensitivity = $`r round(sen,4)`$.  
The test is better to confirm the disease as the specificity is greater than the sensitivity.

a. PPV = $`r round(ppv,4)`$. Thus, we can diagnose the disease with a positive outcome.  
NPV = $`r round(npv,4)`$. Thus, we can rule out the disease with a negative outcome.
:::
:::